---
title: "Activin_Group"
output: html_document
date: '2022-05-03'
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis for RNAseq Elizabeth's Project 

## Logic/ Steps for Analysis 

-  After Initital Analysis with DeSeq2, Tumor vs Benign counts were used to determine Activin Ranking
    - 1st, normalized counts were used from DeSeq2 that were saved in the previous steps
    2nd, genes were pulled from that list and the ranges for the counts were used to calculate the percentiles 
    25th Percentile: 8.4 (Low Activin)
    75th Percentile: 10.9 (High Activin)
    between 25th and 75th Percentile: Intermediate Activin 


```{r message= FALSE, warning= FALSE, echo = FALSE}
library(tximport)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(DESeq2)
```


```{r message= FALSE, warning= FALSE, echo = FALSE}
#tx2gene is formatted to add gene names and required for DESeq2. This only has to be done once for this analysis
txdb <- makeTxDbFromGFF(file="~/Downloads/gencode.v19.annotation.gtf")
saveDb(x=txdb, file = "gencode.v19.annotation.gtf_withproteinids.TxDb")
k <- keys(txdb, keytype = "TXNAME")
tx2gene <- select(txdb, k, "GENEID", "TXNAME")
```

## Group 1: Tumor vs Benign with Activin Grouping 

### Part 1: Setup

```{r message= FALSE, warning= FALSE, echo = FALSE}
dir <- "~/Desktop/quants"
samples <- read.csv(file= "~/Desktop/data_import_benign_vs_tumor.csv")
files <- file.path(dir, samples$SRR, "quant.sf")
all(file.exists(files))
txi <- tximport(files, type = "salmon", tx2gene= tx2gene, ignoreAfterBar= TRUE)
names(txi)
```

### Part 2: DESeq2 for Tumor vs Benign 

```{r echo = FALSE}
library(DESeq2)
sample_meta_data <- read.csv(file= "~/Desktop/Elizabeth_meta_benign_vs_tumor_activin.csv")
```

```{r echo = FALSE}
sampleFiles <- samples$SRR
sampleCondition <- sample_meta_data$Activin
sampleTable <- data.frame(sampleName = sampleFiles,
                          fileName = sampleFiles,
                          condition = sampleCondition)
sampleTable$condition <- factor(sampleTable$condition)
```

```{r echo = FALSE}
colnames(txi$counts) <- rownames(sampleTable)
```

```{r message= FALSE, warning= FALSE, echo = FALSE}
dds <- DESeqDataSetFromTximport(txi, sampleTable, ~condition)
```

```{r message= FALSE, warning= FALSE, echo = FALSE}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r echo = FALSE}
resOrdered <- res[order(res$pvalue),]
```

```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r echo = FALSE}
plotMA(res, ylim=c(-2,2))
```

```{r echo = FALSE}
plotMA(res05, ylim=c(-2,2))
```

### Graphs for Genes of Interest

-   Each circle is a sample in the respective group

-   X axis: group

-   Y axis: Counts

#### INHBA

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000122641.9", intgroup="condition")
```

#### ACVR2A

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000121989.10", intgroup="condition")
```

#### ACVR2B

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000114739.9", intgroup="condition")
```

#### FOXP3

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000049768.10", intgroup="condition")
```

#### p53

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000141510.11", intgroup="condition")
```

#### BRCA1

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000012048.15", intgroup="condition")
```

#### BRCA2

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000139618.10", intgroup="condition")
```

#### INHBB

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000163083.5", intgroup="condition")
```

#### ACVR1B

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000135503.8", intgroup="condition")
```

#### ACVR1C

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000123612.11", intgroup="condition")
```

#### TGFB1

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000105329.5", intgroup="condition")
```


## Group 2: Ascites vs Benign with Activin Grouping 

## Logic/ Steps for Analysis 

-  After Initital Analysis with DeSeq2, Ascites vs Benign counts were used to determine Activin Ranking
    - 1st, normalized counts were used from DeSeq2 that were saved in the previous steps
    2nd, genes were pulled from that list and the ranges for the counts were used to calculate the percentiles 
    25th Percentile: 5.71 (Low Activin)
    75th Percentile: 7.83 (High Activin)
    between 25th and 75th Percentile: Intermediate Activin 

### Part 1: Setup

```{r message= FALSE, warning= FALSE, echo = FALSE}
dir <- "~/Desktop/quants"
samples <- read.csv(file= "~/Desktop/data_import_ascites_vs_normal.csv")
files <- file.path(dir, samples$SRR, "quant.sf")
all(file.exists(files))
txi <- tximport(files, type = "salmon", tx2gene= tx2gene, ignoreAfterBar= TRUE)
names(txi)
```

### Part 2: DESeq2 for Tumor vs Benign 

```{r echo = FALSE}
library(DESeq2)
sample_meta_data <- read.csv(file= "~/Desktop/Elizabeth_meta_benign_vs_ascites_activin.csv")
```

```{r echo = FALSE}
sampleFiles <- samples$SRR
sampleCondition <- sample_meta_data$Activin
sampleTable <- data.frame(sampleName = sampleFiles,
                          fileName = sampleFiles,
                          condition = sampleCondition)
sampleTable$condition <- factor(sampleTable$condition)
```

```{r echo = FALSE}
colnames(txi$counts) <- rownames(sampleTable)
```

```{r message= FALSE, warning= FALSE, echo = FALSE}
dds <- DESeqDataSetFromTximport(txi, sampleTable, ~condition)
```

```{r message= FALSE, warning= FALSE, echo = FALSE}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r echo = FALSE}
resOrdered <- res[order(res$pvalue),]
```

```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r echo = FALSE}
plotMA(res, ylim=c(-2,2))
```

```{r echo = FALSE}
plotMA(res05, ylim=c(-2,2))
```

### Graphs for Genes of Interest

-   Each circle is a sample in the respective group

-   X axis: group

-   Y axis: Counts

#### INHBA

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000122641.9", intgroup="condition")
```

#### ACVR2A

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000121989.10", intgroup="condition")
```

#### ACVR2B

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000114739.9", intgroup="condition")
```

#### FOXP3

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000049768.10", intgroup="condition")
```

#### p53

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000141510.11", intgroup="condition")
```

#### BRCA1

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000012048.15", intgroup="condition")
```

#### BRCA2

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000139618.10", intgroup="condition")
```

#### INHBB

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000163083.5", intgroup="condition")
```

#### ACVR1B

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000135503.8", intgroup="condition")
```

#### ACVR1C

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000123612.11", intgroup="condition")
```

#### TGFB1

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000105329.5", intgroup="condition")
```

