---
title: "30th_Percentile_FOXP3"
output: html_document
date: '2022-05-06'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message= FALSE, warning= FALSE}
library(tximport)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(DESeq2)
library(ggplot2)
```
# Analysis for RNAseq Elizabeth's Project 

## Logic/ Steps for Analysis 

-  After Initial Analysis with DeSeq2, Tumor vs Benign counts were used to determine Activin Ranking
    - 1st, normalized counts were used from DeSeq2 that were saved in the previous steps
    2nd, genes were pulled from that list and the ranges for the counts were used to calculate the percentiles 
    30th Percentile: 5.54 (Low FOXP3)
    70th Percentile:  6.56 (High FOXP3)
    between 30th and 70th Percentile: Intermediate FOXP3
    
    Then only the Ascites and Tumor groups were pulled (no normal samples were in this)


```{r message= FALSE, warning= FALSE, echo = FALSE}
library(tximport)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(DESeq2)
```


```{r message= FALSE, warning= FALSE, echo = FALSE}
#tx2gene is formatted to add gene names and required for DESeq2. This only has to be done once for this analysis
txdb <- makeTxDbFromGFF(file="~/Downloads/gencode.v19.annotation.gtf")
saveDb(x=txdb, file = "gencode.v19.annotation.gtf_withproteinids.TxDb")
k <- keys(txdb, keytype = "TXNAME")
tx2gene <- select(txdb, k, "GENEID", "TXNAME")
```

Eliz## Group 1: Tumor with Activin Grouping 

### Part 1: Setup

```{r message= FALSE, warning= FALSE, echo = FALSE}
dir <- "~/Desktop/quants"
samples <- read.csv(file= "~/Desktop/meta_FoxP3_tumor.csv")
files <- file.path(dir, samples$Sample, "quant.sf")
all(file.exists(files))
txi <- tximport(files, type = "salmon", tx2gene= tx2gene, ignoreAfterBar= TRUE)
names(txi)
```

### Part 2: DESeq2 for Tumor vs Benign 

```{r echo = FALSE}
library(DESeq2)
sample_meta_data <- read.csv(file= "~/Desktop/meta_FoxP3_tumor.csv")
```

```{r echo = FALSE}
sampleFiles <- samples$Sample
sampleCondition <- sample_meta_data$FOXP3
sampleTable <- data.frame(sampleName = sampleFiles,
                          fileName = sampleFiles,
                          condition = sampleCondition)
sampleTable$condition <- factor(sampleTable$condition)
```

```{r echo = FALSE}
colnames(txi$counts) <- rownames(sampleTable)
```

```{r message= FALSE, warning= FALSE, echo = FALSE}
dds <- DESeqDataSetFromTximport(txi, sampleTable, ~condition)
```

```{r message= FALSE, warning= FALSE, echo = FALSE}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r echo = FALSE}
resOrdered <- res[order(res$pvalue),]

write.csv(x=resOrdered, file= "~/Desktop/FOXP3_30th_70th_percentile.csv")
```


```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r echo = FALSE}
plotMA(res, ylim=c(-2,2))
```

```{r echo = FALSE}
plotMA(res05, ylim=c(-2,2))
```

### Graphs for Genes of Interest

-   Each circle is a sample in the respective group

-   X axis: group

-   Y axis: Counts

#### INHBA

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000122641.9", intgroup="condition")
```

#### ACVR2A

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000121989.10", intgroup="condition")
```

#### ACVR2B

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000114739.9", intgroup="condition")
```

#### FOXP3

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000049768.10", intgroup="condition")
```

#### INHBB

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000163083.5", intgroup="condition")
```

#### ACVR1B

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000135503.8", intgroup="condition")
```

#### ACVR1C

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000123612.11", intgroup="condition")
```

#### TGFB1

```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000105329.5", intgroup="condition")
```

#### CD8A
```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000153563.11", intgroup="condition")
```

#### CD3G
```{r echo = FALSE}
plotCounts(dds, gene= "ENSG00000160654.5", intgroup="condition")
```

```{r}
vsd <- vst(dds, blind=FALSE)
plotPCA(vsd, intgroup=c("condition"))
```

```{r}
ntd <- normTransform(dds)
library("vsn")
meanSdPlot(assay(ntd))
```


```{r}
library("pheatmap")
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(dds)[,c("condition")])
pheatmap(assay(ntd)[select,], cluster_rows=FALSE, show_rownames=TRUE,
         cluster_cols=TRUE, annotation_col=df)
```
The 2 genes that how a difference in expression are IGKC and IGHG1


